---
title: "Homework_1"
author: "Sai Saketh Boyanapalli"
date: "August 29, 2017"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Packages required
```{r}
library(moments)
library(plyr)
library(datasets)
```

## Question 1

# 1 Using R: Vectors
(a) Create a vector with 10 numbers (3, 12, 6, -5, 0, 8, 15, 1, -10, 7) and assign it to x.
```{r}
x <- c(3, 12, 6, -5, 0, 8, 15, 1, -10, 7)
x
```

(b) Using the commands seq, min, and max with one line of code create a new vector y with 10 elements
ranging from the minimum value of x to the maximum value of x.
``` {r}
y <- seq(min(x),max(x),length.out = 10)
y
```

(c) Compute the sum, mean, standard deviation, variance, mean absolute deviation, quartiles, and
quintiles for x and y.
```{r}

#sum of x and y
sum(x)
sum(y)

#mean of x and y
mean(x)
mean(y)

#standard deviation of x and y
sd(x)
sd(y)

#variance of x and y
var(x)
var(y)

#mean absolute deviation of x and y
mad(x)
mad(y)

#quartiles of x and y
quantile(x)
quantile(y)

#quintiles of x and y
quantile(x,probs = seq(0,1,0.2))
quantile(y,probs = seq(0,1,0.2))

```

(d) Create a new 7 element vector z by using R to randomly sample from x with replacement.
```{r}
z<- sample(x,7, replace = TRUE)
z
```
(e) Find a package (or packages) that provide the statistical measures skewness and kurtosis. Use the
appropriate functions from the package to calculate the skewness and kurtosis of x.

The Skewness and Kurtosis functions are available in moments package.
```{r}
skewness(x)
kurtosis(x)
```

(f) Use t.test() to compute a statistical test for differences in means between the vectors x and y.
Are the differences in means signifcant?
```{r}
t.test(x,y)
```

(g) Sort the vector x and re-run the t-test as a paired t-test.
```{r}
sorted_x <- sort(x)
t.test(sorted_x,y,paired = TRUE)
```
(h) Create a logical vector that identifies which numbers in x are negative.
```{r}
neq_x <- x[x<0]
```
(i) Use this logical vector to remove all entries with negative numbers from x. (Make sure to overwrite
the vector x so that the new vector x has 8 elements!)
```{r}
x <- x[!x %in% neq_x]
x
```

##Question 2
#Using R: Introductory data exploration

This exercise relates to the College data set, which can be found in the file \College.csv" in D2L. The
file contains a number of variables for 777 different universities and colleges in the US.

#(a) 
Use the read.csv() function to read the data into a data frame in R. Call the data frame college.
Make sure that you have the directory set to the correct location for the data (or that the data is
in the same directory as the RStudio project).
```{r}
college <- read.csv("college.csv",header = TRUE)
head(college)
```
#(b) 

```{r}
# this will assisgn the row names of the data frame to college names
rownames (college) <- college [,1]
View (college )
head(college)

# Now that we have assigned each row to the appropriate college name we can remove the column with college names
college <- college [,-1]
head(college)
```
#(c)
(i)
summary() function will give us the summary of the data
```{r}
summary(college)
```
(ii)
```{r}
?pairs # using ? before a function shows us the documentation for it.
pairs(college[,1:10],college)
```
(iii)
```{r}
# This function creates a boxplot for no of OutState students in private colleges
plot(college$Private,college$Outstate,main = "OutState students in Private College", xlab = "Private",ylab = "Outstate" )
```
iv. Using the following bit of code you will create a new qualitative variable, called Elite by
binning the Top10perc variable. That is, Elite will classify the universities into two groups
based on whether or not the proportion of students coming from the top 10% of their high
school classes exceeds 50%.
Add comments to each line below explaining what the corresponding code is doing and then
run the code.
```{r}
Elite <- rep ("No", nrow(college )) # this line creates a list with value 'NO' with the length set to no of rows in college. using rep function. 
Elite [college$Top10perc >50] <- "Yes" # In this line the college with top10percent greater than 50, the elite value is set to "Yes"
Elite <- as.factor (Elite) #The values in Elite are factored to two levels    
college <- data.frame(college ,Elite) # Elite is addes as one of the variables to college data frame.
```

v. Use the summary() function to see how many elite universities there are.
```{r}
summary(college$Elite)
```
we can see there are 78 elite students in total.
vi.
```{r}
plot(college$Elite,college$Outstate,main = "Outstate elite students", xlab = "Elite students", ylab = "OutState")
```
we can clearly see that there are more outstate elite students.

vii. Use the hist() function to produce some histograms with diering numbers of bins for a
few of the quantitative variables.
```{r}
par(mfrow=c(2,2)) # this command will divide plot window into 4 sections
hist(college$Outstate,breaks = 10) # this will create a Histogram.
hist(college$Apps,breaks = 5) # breaks is used to set no of bins.
hist(college$Accept,breaks = 15)
hist(college$Top10perc,breaks = 6)

```
##Question 3 
#Using R: Manipulating data in data frames
(a) Load the data frame baseball in the plyr package. Use ?baseball to get information about the
data set and definitions for the variables.
```{r}
data("baseball") # data is used to load a specific data set
?baseball

```
(b) You will calculate the on base percentage for each player, but first clean up the data:
* Before 1954, sacrifice flies were counted as part of sacrifice hits, so for players before 1954,sacrifice flies (i.e. the variable sf) should be set to 0.
```{r}
baseball$sf[baseball$year < 1954] <- 0
```

* Hit by pitch (the variable hbp) is often missing { set these missings to 0.
```{r}
baseball$hbp[is.na(baseball$hpb)] <- 0
```

* Exclude all player records with fewer than 50 at bats (the variable ab).
```{r}
baseball <- baseball[-c(baseball$ab < 50), ]
```

(c) Compute on base percentage in the variable obp according to the formula:
```{r}
obp <-((baseball$h + baseball$bb + baseball$hbp)/(baseball$ab + baseball$bb + baseball$hbp + baseball$sf)) 
baseball <- data.frame(baseball,obp)
```
(d) Sort the data based on the computed obp and print the year, player name, and on base percentage for the top five records based on this value.
```{r}
Sorted_obp <- baseball[order(-obp) , ] # (-obp indicates decreasing order)
top_five <- Sorted_obp[1:5, ]
top_five[,c("year","id","obp")]
```
##Question 4 
#Using R: aggregate() function
(a) Load the quakes data from the datasets package.
```{r}
data("quakes")
```
(b) Plot the recorded earthquake magnitude against the earthquake depth using the plot command.
```{r}
plot(quakes$depth,quakes$mag,main ="earthquake magnitude against the earthquake depth", xlab = "depth",ylab = "magnitude")
```
(c) Use aggregate to compute the average earthquake depth for each magnitude level. Store these results in a new data frame named quakeAvgDepth.
```{r}
quakeAvgDepth <- aggregate(quakes$depth ~ quakes$mag,quakes, FUN = mean)
```
(d) Rename the variables in quakeAvgDepth to something meaningful.
```{r}
colnames(quakeAvgDepth) <- c("Magnitude of Earthquake","corresponding_Average_Depth")
head(quakeAvgDepth)
```
(e) Plot the magnitude vs. the average depth.
```{r}
plot(quakeAvgDepth$corresponding_Average_Depth,quakeAvgDepth$`Magnitude of Earthquake`,main="Magnitude vs. the Average Depth of Quake",xlab="Average Depth",ylab="Magnitude")
```
(f) From the two plots, do you think there is a relationship between earthquake depth and magnitude?

From the Two graphs we can see that the depth of the quake decreases with the increase in magnitude.

